Constraining gravitational wave amplitude birefringence with GWTC-3

### 2022-12-08 src/tex/ms.tex
Updated content

\subsection{Parameter estimation}
% Parameter estimation
To constrain $\kappa$, we perform PE with data from the third GW transient catalog \citep{GWTC-2.1, GWTC-3}, GWTC-3, using Bilby \citep{Bilby}, a Bayesian toolkit for GW data analysis which can calculate posteriors of GW parameters based on interferometer data and priors for the parameters in question.
We modified Bilby to include the birefringence effect in the likelihood function.
We then perform PE on intrinsic and extrinsic parameters including $\kappa$ with open data from the Gravitational Wave Open Science Center \citep{GWOSC}.
For PE configuration, we use the same PE configuration as in \citet{GWTC-2.1, GWTC-3} for each event and set the prior on $\kappa$ to a uniform distribution between $-1$ and $1$.
We also perform a sanity test to check if our PE could recover the PE results done by LVK in \citet{GWTC-2.1, GWTC-3}.
To implement the sanity check, we need to assume there is no birefringence by setting $\kappa$ to be $0$, which forces the PE results to be consistent with GR.
These sanity tests result can be found in \citet{dataset}.

@@

%cite Max's paper

% under XXX, kappa should share the same value
% to understand more, the true value of kappa is drawn from a distribution
% the distribution should be consistent with a delta function
% if it is not a delta function, there should be more study (model not fully correct, new physics, or systematic error)

% we expect the kappa to be a single value for all events at the beginning
With the PE results of the individual events, we can perform Bayesian hierarchical modeling to constrain the birefringence effect.
We treated $\kappa$ posteriors from all events as a set of measured $\kappa$ from an underlying distribution.
% instead of assuming all event share one value, we assume the kappa is a true value from a distribution.
The underlying distribution can be assumed to be a Dirac delta function, which means there is a single true value of $\kappa$, or other distributions, which means there is a range of plausible values of $\kappa$.
% we expect the delta should be 
In this study, we assume the underlying distribution is a Gaussian distribution with mean $\mu$ and standard deviation $\sigma$.
% Gaussian max entropy, whatever the true distribution is, the mean and the standard deviation will be true.
With Bayesian hierarchical modeling, which is a method to combine multiple measurements of a random variable by assuming the measurements are independent and identically distributed, we can calculate the posterior of $\mu$ and $\sigma$ from the set of measured $\kappa$ posterior.
The posterior of $\mu$ and $\sigma$ can be calculated by
\begin{equation}
    P(\mu,\sigma|\{d_i\})\propto P(\mu,\sigma)\prod_{i}\int\frac{P(\kappa_i|d_i)}{P(\kappa_i}P(\kappa_i|\mu,\sigma)\,d\kappa_i\,,
    \label{eq:posterior_of_mu_sigma}
\end{equation}
where function $P$ is the probability density function, and $\{d_i\}$ is the set of GW event data.
Note that $P(\kappa)$ is the prior on $\kappa$, which is a uniform distribution between $-1$ and $1$ in this study.
We chose the prior of $\mu$ and $\sigma$ being uniform distributions.
Equation \ref{eq:posterior_of_mu_sigma} can then be simplified to
\begin{equation}
    P(\mu,\sigma|\{d_i\})\propto\prod_{i}\int P(\kappa_i|d_i)\mathcal{N}(\mu,\sigma)\,d\kappa_i\,.
\end{equation}
To sample the posterior distribution of $\mu$ and $\sigma$, we used a generic sampling package, flowMC \citep{flowMC}.
We then calculate the population constraint on $\kappa$ from the posterior of $\mu$ and $\sigma$.
The constraint was calculated by
\begin{equation}
    P(\kappa|\{d_i\})=\int \mathcal{N}(\mu,\sigma)P(\mu,\sigma|\{d_i\})\,d\mu\,d\sigma\,.
\end{equation}
% np.choose and weight

@@

>>>>

\subsection{Parameter estimation}
% Parameter estimation
To constrain $\kappa$, we perform PE with data from the third GW transient catalog \citep{GWTC-2.1, GWTC-3}, GWTC-3, using Bilby \citep{Bilby}, a Bayesian toolkit for GW data analysis which can calculate posteriors of GW parameters based on interferometer data and priors for the parameters in question.
We modified Bilby to include the birefringence effect in the likelihood function.
We then perform PE on intrinsic and extrinsic parameters including $\kappa$ with open data from the Gravitational Wave Open Science Center \citep{GWOSC}.
For PE configuration, we use the same PE configuration as in \citet{GWTC-2.1, GWTC-3} for each event and set the prior on $\kappa$ to a uniform distribution between $-1$ and $1$.
The PE results are published in \citet{dataset}.

We assume $\kappa$ is the same for all events and calculate a posterior distribution of $\kappa$.
The posterior distribution is given by
\begin{equation}
    p(\kappa|\{d\})\propto \prod{i}p(\kappa|d_i)\,,
    \label{eq:restricted_posterior}
\end{equation}
where $p$ is the probability density function, $i$ is the index of the event, $\{d\}$ is the set of data including all events, and $d_i$ is the data of a single event.
Equation \ref{eq:restricted_posterior} shows that the posterior distribution of $\kappa$ is the product of the posterior distribution of $\kappa$ for each event.
This posterior distribution is a restricted posterior distribution because we assume $\kappa$ is the same for all events.

Note that we also perform a sanity test to check if our PE could recover the PE results done by LVK in \citet{GWTC-2.1, GWTC-3}.
To implement the sanity check, we set $\kappa$ to be $0$, which forces the PE results to be consistent with GR.
These sanity test results are also attached in \citet{dataset}.

@@

We then perform Bayesian hierarchical modeling to further analyze the constraint on $\kappa$.
Bayesian hierarchical modeling is a method to check if our waveform model can fully describe the amplitude birefringence effect.
Here, instead of assuming all events share the same value of $\kappa$, we assume $\kappa$ from each event is a true value drawn from a Gaussian distribution with mean $\mu$ and standard deviation $\sigma$.
We chose a Gaussian distribution as the underlying distribution of $\kappa$ because it captures the correct mean and the standard deviation of the distribution no matter what the underlying distribution is.

With the posterior of $\kappa$ from PE on each event, the posterior of $\mu$ and $\sigma$ can be calculated by
\begin{equation}
    p(\mu,\sigma|\{d\})\propto p(\mu,\sigma)\prod_{i}\int\frac{p(\kappa_i|d_i)}{p(\kappa_i}p(\kappa_i|\mu,\sigma)\,d\kappa_i\,,
    \label{eq:posterior_of_mu_sigma}
\end{equation}
Note that $p(\kappa)$ is the prior of $\kappa$, which is a uniform distribution between $-1$ and $1$ in this study.
We chose the prior of $\mu$ and $\sigma$ being uniform distributions.
Equation \ref{eq:posterior_of_mu_sigma} can then be simplified to
\begin{equation}
    p(\mu,\sigma|\{d_i\})\propto\prod_{i}\int p(\kappa_i|d_i)\mathcal{N}(\mu,\sigma)\,d\kappa_i\,.
\end{equation}
To sample the posterior distribution of $\mu$ and $\sigma$, we used a sampling package, flowMC \citep{flowMC}.

If our model can fully describe the birefringence effect, the true value of $\kappa$ should be a single value, and the Gaussian distribution should have a standard deviation of $0$.
If the standard deviation obtained is not $0$, it means our model cannot fully describe the birefringence effect.
It could be due to our model does not fully capture the physics of birefringence, or there is a systematic error in the data.

We then calculate the generic posterior distribution of $\kappa$ from the posterior of $\mu$ and $\sigma$.
The constraint was calculated by
\begin{equation}
    p(\kappa|\{d_i\})=\int \mathcal{N}(\mu,\sigma)p(\mu,\sigma|\{d_i\})\,d\mu\,d\sigma\,.
\end{equation}

@@

% np.choose and weight for drawing samples
